---
title: "Assignment1 R-Basics"
output:
html_document: default
autor: Sonal Savaliya
---


**1.** 
1. Suppose you keep track of your mileage each time you fill up your car's gas tank. At your last fill-ups, the
mileage values were : 65311 , 65624 , 65908 , 66219 , 66499 , 66821 , 67145 , 67447 respectively

**a.** Enter these numbers into an R vector. Use the function diff() on the data. What is the result?

```{r}
mileage_values = c(65311 , 65624 , 65908 , 66219 , 66499 , 66821 , 67145 , 67447)
x = diff(mileage_values)
print(x)

## 313 284 311 280 322 324 302
```
Based on the result above, I think the diff() function gives the difference of consecutive numbers of a vector.


**b.** Use the function length() to find the lengths of the miles and x vectors. Explain the result.

```{r}
length(mileage_values) # length of vector
## 8

length(x)
## 7
```

The results suggest that length function returns the length of a vector.

**c.** Use R functions to find the maximum and minimum number of miles between fill-ups, the average number of
miles between fill-ups, the standard deviation. Hint: Refer Statistical functions in CheatSheet

```{r}
max(x)
## 324

min(x)
## 280

mean(x)
## 305.1429

sd(x)
## 17.43969
```


**2.** You have tracked your commute times for two weeks (10 days) and you recorded the following times in minutes:
17 16 20 24 22 15 21 15 17 22
    
**a.**  Enter this into a vector called commute. Use R functions to find the average and a statistical summary.

```{r}
commute = c(17,16,20,24,22,15,21,15,17,22)
mean(commute)
## 18.9

summary(commute)
##   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
##  15.00   16.25   18.50   18.90   21.75   24.00 
```

**b.** Suppose the 24 was a mistake; it should have been 18. How can you fix this? Do so, and then find the new
average. Hint: Index into the vector, then use an assignment statement to fix this mistake.

```{r}
commute[4] = 18  #here, commute's 4th value will be change 24 to 18
mean(commute)
## 18.3
```

**c.** How many times was your commute 20 minutes or more?

```{r}
sum(commute >= 20)
## 4

# OR
length(which(commute >= 20))
## 4
```

**d.**  What percent of your commutes are less than 17 minutes? How can you answer this with R?

```{r}
#30% of commutes are less than 17 minutes.
commute_lessthan_17_per = sum(commute < 17)/length(commute) * 100
print(commute_lessthan_17_per)
## 30
```

**e.** Briefly explain the result of the following commands. Hint: Refer Vector data extraction summary in
CheatSheet
- commute + 1
- commute[-6]
- sum(commute>21 | commute<16)
- commute[commute>=17]
- sum(commute[commute>20])
- commute[4:6]

```{r}
commute + 1
## 18 17 21 19 23 16 22 16 18 23
```
The function above adds one to (I almost answer it for you here :-) each element of the vector.

```{r}
commute[-6]
## 17 16 20 18 22 21 15 17 22
```
The function above removes 6th element of the vector.

```{r}
sum(commute>21 | commute<16)
## 4
```
The function above returns a set of elements that are greater than 21 and less than 16 from the provided vector. It does not give any elements between 16 to 21 includig 16 and 21. 

```{r}
commute[commute>=17]
## 17 20 18 22 21 17 22
```
The function above extracts all elements which are greater than and equal to 17 from the vector. 

```{r}
sum(commute[commute>20])
## 65
```
The function above gives sum of all values which are greater than 20 from the vector.
```{r}
commute[4:6]
## 18 22 15
```
The function above gives all element values from 4 to 6 index of the vector.


**3.** Let the data x be given by

```{r}
x = c(1, 8, 2, 6, 3, 8, 5, 5, 5, 5) 

#3.a.

```
**3.a** (X1 + X2 + X3. . .+X10)/10 use sum

```{r}
sum(x)/10
##  4.8
```

**3.b** Find log10(Xi) for each i. (Use the log function which by default is base e)

```{r}
log10(x)
## [1] 0.0000000 0.9030900 0.3010300 0.7781513 0.4771213 0.9030900 0.6989700 0.6989700 0.6989700
## [10] 0.6989700

```

**3.c** Find (X4.4 )/2.875 for each i. (Do it all at once)
```{r}
x^4.4/2.875
## [1]    0.3478261 3273.0911040    7.3433484  923.0551563   43.7215623 3273.0911040  413.8378128
## [8]  413.8378128  413.8378128  413.8378128

```

**3.d** Find the difference between the largest and smallest values of x. (This is the range. You can use max and
min or guess a built in command.)

```{r}
diff_large_small = max(x) - min(x)
print(diff_large_small)
## 7
```

**4.** Your cell phone bill varies from month to month. Suppose that your bills had the following monthly amounts last
year:
```{r}
bill = c(46, 33, 39, 37, 46, 30, 48, 32, 49, 35, 30, 48)

```

  **a.**  Run the following calculations, and explain its result: range(bill) range(bill)[2] - range(bill)[1]
  
```{r}
range(bill)
##  30 49

range(bill)[2] - range(bill)[1]
## 19
```
The function above range(bill) gives the maximim and minimum of the given vector.
The function range(bill)[2]-range(bill)[1] gives the difference between the largest and the smallest cell phone bill.


  **b.** How many months was the amount greater than $40? What percentage was this?
  
```{r}
months_greaterthan_40 = sum(bill > 40)
print(months_greaterthan_40)
## 5

bill_greaterthan_40_per = sum(bill > 40)/length(bill) * 100
print(bill_greaterthan_40_per)
## 41.66667
```

